---
title: "PS1_Shiyun"
output: pdf_document
date: "2025-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Problem set 1 

## Q1 
1)	Calculate the 10-year annualized real excess returns. What is the mean excess return since 1881? Excess returns is defined as the difference between stock returns and bond returns.
```{r}
library(readxl)
Shiller_dataset <- read_excel("~/Desktop/behavioral finance /HW 1 20250131/Shiller_dataset.xlsx")
View(Shiller_dataset)
head(Shiller_dataset)
```
# new column for excess return = stock real return - bond real return
```{r}
Shiller_dataset$`10_yr_real_excess_ret`<- Shiller_dataset$`10_yr_ann_stock_real_return`-Shiller_dataset$`10_yr_ann_bond_real_ret`
```
#Ans: shown in excel sheet

#mean excess return, select date range post 1881, and remove na
```{r}
mean(Shiller_dataset[Shiller_dataset$date >= 1881.01,]$`10_yr_real_excess_ret`, na.rm = T)
```
#Ans: [1] 0.04584711
## Q2 
2)	Calculate the excess CAPE yield (Defined as 1/ CAPE – 10-year real treasury yield). What is the mean excess CAPE yield and its standard deviation since 1881? 

## excess CAPE yield = 1/CAPE - 10 year real treasury 
```{r}
Shiller_dataset$excess_CAPE_yield <- 1 / Shiller_dataset$CAPE - Shiller_dataset$`10_yr_treasury_real_yield`
```
#Ans:show in excel sheet

##mean excess CAPE yield, select data range post 1881, and remove na 
```{r}
mean_excess_CAPE_yield<-mean(Shiller_dataset[Shiller_dataset$date >= 1881.01,]$`excess_CAPE_yield`, na.rm = T)
mean_excess_CAPE_yield
```
##Ans:[1] 0.04621346

## Q3 
3)	Plot the excess CAPE yield and subsequent 10-year real excess returns since 1881, i.e. replicate the plot on slide 4 of the first lecture. 
```{r}
install.packages("ggplot2") #install package
library(ggplot2) #call package

#transform numeric date to Ymd format 
Shiller_dataset$date <- as.Date(paste0(Shiller_dataset$date, ".01"), format = "%Y.%m.%d")
# x-axis is date, compute line plot, y-axis is 10 yr real excess return
ggplot(data = Shiller_dataset, aes(x = date)) + 
  geom_line(aes(y = `10_yr_real_excess_ret`), color = "black") +
   geom_line(aes(y = `excess_CAPE_yield`), color = "red") +
  ylab("10 year real excess return & excess CAPE yield")  +
  xlab("year")
```
## Q4 
4)	Run a linear regression of 10-year real excess returns on excess CAPE yield. What is the regression coefficient and the p-value? Interpret these values. 

```{r}
model1 <- lm(`10_yr_real_excess_ret` ~ `excess_CAPE_yield`, Shiller_dataset)
summary(model1)
```
#Ans: The coefficient is 0.584034, with a p-value of 2e-16. Based on the regression coefficient, this means that for each 1unit increase in excess_cape_yield, the 10_yr_real_excess_return increases by approximately 0.584 units. The p-value in this case is extremely small, indicate the relationship between excess cape yield and 10 year real excess return is highly statistically signigicant. We can reject the null hypothesis. Overall, these two values suggest a strong and statistically significant relationship between two variables. 
